---
title: "assign3"
author: "Ryan Huang"
date: "3/23/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(tidyverse)
library(rethinking)
library(purrr)
library(rlist)
```
### Assignment 3
# Data preprocessing
```{r}
data(foxes)
summary(foxes)
data_assign3 = foxes
data_assign3$area_z = as.vector(scale(data_assign3$area))
data_assign3$avgfood_z = as.vector(scale(data_assign3$avgfood))
data_assign3$groupsize_z = as.vector(scale(data_assign3$groupsize))
plot(data_assign3$area_z, data_assign3$weight)
hist(data_assign3$weight)
```
## Question1
# 1-1 Body weight as a linear function of territory size (area)
```{r}
m1.1_draft = "
data {
  int N;
  vector[N] weight;
  vector[N] area;
}
parameters {
  real alpha;
  real betaA;
  real sigma;
}
model {
  vector[N] mu = alpha + betaA * area;
  
  weight ~ normal(mu, sigma);
  alpha ~ normal(4.5,1);
  betaA ~ normal(0,5);
  sigma ~ uniform(0,5);
}
"
m1.1_data = list(
  N = nrow(data_assign3),
  weight = data_assign3$weight,
  area = data_assign3$area_z
)
m1.1_fit = stan(model_code = m1.1_draft, data = m1.1_data, iter = 1000, cores = 4, chains = 2)
summary(m1.1_fit)
plot(m1.1_fit, plotfun = "trace", pars = c("alpha", "betaA","sigma"), inc_warmup = TRUE)

```
# 1-2 Body weight as a linear function of groupsize
```{r}
m1.2_draft = "
data {
  int N;
  vector[N] weight;
  vector[N] groupsize;
}
parameters {
  real alpha;
  real betaG;
  real sigma;
}
model {
  vector[N] mu = alpha + betaG * groupsize;
  
  weight ~ normal(mu, sigma);
  alpha ~ normal(4.5,1);
  betaG ~ normal(0,5);
  sigma ~ uniform(0,5);
}
"
m1.2_data = list(
  N = nrow(data_assign3),
  weight = data_assign3$weight,
  groupsize = data_assign3$groupsize_z
)
m1.2_fit = stan(model_code = m1.2_draft, data = m1.2_data, iter = 1000, cores = 4, chains = 2)
plot(m1.2_fit, plotfun = "trace", pars = c("alpha", "betaG","sigma"), inc_warmup = TRUE)
summary(m1.2_fit)
```
# 1-3 Plot predicted regression line and the 95% interval of the mean.
```{r}
interval_est = function(alpha,beta, x, bound, interval = 0.89){
  mu_at_x = alpha + beta * x
  if (bound == 'lower'){
    return(HPDI(mu_at_x, prob = interval)[1])
  } else if (bound == 'upper'){
    return(HPDI(mu_at_x, prob = interval)[2])
  } else {
    return(mean(mu_at_x))
  }
}
```
# m1.1
```{r}
# m1.1
m1.1_result = as.data.frame(m1.1_fit)
new_area = seq(1,5, length.out = 20)
m1.1_prediction = 
  tibble(area = new_area,
         predict_weight = purrr::map(new_area - mean(data_assign3$area),interval_est,
                                     alpha = m1.1_result$alpha,
                                     beta = m1.1_result$betaA,
                                     bound='')%>% unlist() %>% as.vector(),
         upper_95       = purrr::map(new_area - mean(data_assign3$area),interval_est,
                                     alpha = m1.1_result$alpha,
                                     beta = m1.1_result$betaA,
                                     bound ='upper',
                                     interval = 0.95)%>% unlist() %>% as.vector(),
         lower_95       = purrr::map(new_area - mean(data_assign3$area),interval_est,
                                     alpha = m1.1_result$alpha,
                                     beta = m1.1_result$betaA,
                                     bound='lower',
                                     interval = 0.95)%>% unlist() %>% as.vector()
         ) 
m1.1_plot = ggplot() + 
  geom_line(data = m1.1_prediction,
    aes(x = area, y = predict_weight)) + 
  geom_ribbon(data = m1.1_prediction,
              aes(x = area, ymin = lower_95, ymax = upper_95),
              alpha = .5)
```
```{r}
# From the following plot, we notice that area is not highly important.
# Also from the summary report, the slope of betaArea is very small.
m1.1_plot
```

#m1.2
```{r}
m1.2_result = as.data.frame(m1.2_fit)
new_groupsize = seq(1,8, length.out = 20)
m1.2_prediction = 
  tibble(groupsize = new_groupsize,
         predict_weight = purrr::map(new_groupsize - mean(data_assign3$groupsize),interval_est,
                                     alpha = m1.2_result$alpha,
                                     beta = m1.2_result$betaG,
                                     bound='') %>% unlist() %>% as.vector(),
         upper_95       = purrr::map(new_groupsize - mean(data_assign3$groupsize),interval_est,
                                     alpha = m1.2_result$alpha,
                                     beta = m1.2_result$betaG,
                                     bound ='upper',
                                     interval = 0.95) %>% unlist() %>% as.vector(),
         lower_95       = purrr::map(new_groupsize - mean(data_assign3$groupsize),interval_est,
                                     alpha = m1.2_result$alpha,
                                     beta = m1.2_result$betaG,
                                     bound='lower',
                                     interval = 0.95) %>% unlist() %>% as.vector()
         )

m1.2_plot = ggplot() + 
  geom_line(data = m1.2_prediction,
    aes(x = groupsize, y = predict_weight)) + 
  geom_ribbon(data = m1.2_prediction,
              aes(x = groupsize, ymin = lower_95, ymax = upper_95),
              alpha = .5)
```
```{r}
# From the following plot, we notice that groupsize is important.
# Also from the summary report, the slope for beta groupsize is different from 0.
m1.2_plot
```

## Question 2 
# 2-1 Fit weight = alpha + area + groupsize
```{r}
m2.1_draft = "
data {
  int N;
  vector[N] area;
  vector[N] groupsize;
  vector[N] weight;
}
parameters {
  real alpha;
  real betaA;
  real betaG;
  real sigma;
}
model{
  vector[N] mu = alpha + betaA * area + betaG * groupsize;
  
  weight ~ normal(mu, sigma);
  alpha ~ normal(4.5, 1);
  betaA ~ normal(0, 5);
  betaG ~ normal(0, 5);
  sigma ~ uniform(0,5);
}
"
m2.1_data = list(
  N = nrow(data_assign3),
  groupsize = data_assign3$groupsize_z,
  area = data_assign3$area_z,
  weight = data_assign3$weight
)
m2.1_fit = stan(model_code = m2.1_draft, data = m2.1_data, iter = 1000,  chains = 2, cores = 4)

plot(m2.1_fit, plotfun = "trace", pars = c("alpha", "betaG","sigma"), inc_warmup = TRUE)
summary(m2.1_fit)
```
# 2-2 Counterfactual plot for area
```{r}
interval_est_2 = function(alpha, beta1, beta2, x1, x2, bound, interval = 0.89){
  mu_at_x = alpha + beta1 * x1 + beta2 * x2
  if (bound == 'lower'){
    return(HPDI(mu_at_x, prob = interval)[1])
  } else if (bound == 'upper'){
    return(HPDI(mu_at_x, prob = interval)[2])
  } else {
    return(mean(mu_at_x))
  }
}
```

```{r}
# Fixed groupsize
m2.1_result = as.data.frame(m2.1_fit)
new_area = seq(1,5, length.out = 20)
m2.1.1_prediction = 
  tibble(area = new_area,
         predict_weight = purrr::map(new_area - mean(data_assign3$area),interval_est_2,
                                     alpha = m2.1_result$alpha,
                                     beta1 = m2.1_result$betaA,
                                     beta2 = m2.1_result$betaG,
                                     x2 = 0,
                                     bound='')%>% unlist() %>% as.vector(),
         upper_95       = purrr::map(new_area - mean(data_assign3$area),interval_est_2,
                                     alpha = m2.1_result$alpha,
                                     beta1 = m2.1_result$betaA,
                                     beta2 = m2.1_result$betaG,
                                     x2 = 0,
                                     bound ='upper',
                                     interval = 0.95)%>% unlist() %>% as.vector(),
         lower_95       = purrr::map(new_area - mean(data_assign3$area),interval_est_2,
                                     alpha = m2.1_result$alpha,
                                     beta1 = m2.1_result$betaA,
                                    beta2 = m2.1_result$betaG,
                                     x2 = 0,
                                     bound='lower',
                                     interval = 0.95)%>% unlist() %>% as.vector()
         ) 
m2.1.1_plot = ggplot() + 
  geom_line(data = m2.1.1_prediction,
    aes(x = area, y = predict_weight)) + 
  geom_ribbon(data = m2.1.1_prediction,
              aes(x = area, ymin = lower_95, ymax = upper_95),
              alpha = .5)
m2.1.1_plot
```
# 2-3 Counterfactual plot for group size
```{r}
new_groupsize = seq(1,8, length.out = 20)
m2.1.2_prediction = 
  tibble(groupsize = new_groupsize,
         predict_weight = purrr::map(new_groupsize - mean(data_assign3$groupsize),interval_est_2,
                                     alpha = m2.1_result$alpha,
                                     beta1 = m2.1_result$betaA,
                                     beta2 = m2.1_result$betaG,
                                     x1 = 0,
                                     bound='') %>% unlist() %>% as.vector(),
         upper_95       = purrr::map(new_groupsize - mean(data_assign3$groupsize),interval_est_2,
                                     alpha = m2.1_result$alpha,
                                     beta1 = m2.1_result$betaA,
                                     beta2 = m2.1_result$betaG,
                                     x1 = 0,
                                     bound ='upper',
                                     interval = 0.95) %>% unlist() %>% as.vector(),
         lower_95       = purrr::map(new_groupsize - mean(data_assign3$groupsize),interval_est_2,
                                     alpha = m2.1_result$alpha,
                                     beta1 = m2.1_result$betaA,
                                     beta2 = m2.1_result$betaG,
                                     x1 = 0,
                                     bound='lower',
                                     interval = 0.95) %>% unlist() %>% as.vector()
         )

m2.1.2_plot = ggplot() + 
  geom_line(data = m2.1.2_prediction,
    aes(x = groupsize, y = predict_weight)) + 
  geom_ribbon(data = m2.1.2_prediction,
              aes(x = groupsize, ymin = lower_95, ymax = upper_95),
              alpha = .5)
m2.1.2_plot
```
# 2-4 What does this model say about the importance of each variable? Why do you get different results than you got in the questions just above?

From the 2 above plots, we can notice that the 2 variables are both important in the multivaritive model. The reason why the results are different from that of Q1 is that the effect from area is masked in the bivaritive model.

## Question 3
# 3-1 body weight as an additive function of avgfood and groupsize
```{r}
m3.1_draft = "
data {
  int N;
  vector[N] avgfood;
  vector[N] groupsize;
  vector[N] weight;
}
parameters {
  real alpha;
  real betaAF;
  real betaG;
  real sigma;
}
model {
  vector[N] mu = alpha + betaAF * avgfood + betaG * groupsize;

  weight ~ normal(mu, sigma);
  alpha ~ normal(4.5, 1);
  betaAF ~ normal(0, 5);
  betaG ~ normal(0, 5);
  sigma ~ uniform(0,5);
}
"
m3.1_data = list(
  N = nrow(data_assign3),
  avgfood = data_assign3$avgfood_z,
  groupsize = data_assign3$groupsize_z,
  weight = data_assign3$weight
)

m3.1_fit = stan(model_code = m3.1_draft, data = m3.1_data, iter = 1000, cores = 4, chains = 2)
plot(m3.1_fit, plotfun = "trace", pars = c("alpha", "betaG","betaAF","sigma"), inc_warmup = TRUE)
summary(m3.1_fit)
```
# 3-2 body weight as an additive function of, avgfood and groupsize and area
```{r}
m3.2_draft = "
data {
  int N;
  vector[N] avgfood;
  vector[N] groupsize;
  vector[N] weight;
  vector[N] area;
}
parameters {
  real alpha;
  real betaAF;
  real betaG;
  real betaA;
  real sigma;
}
model {
  vector[N] mu = alpha + betaAF * avgfood + betaG * groupsize + betaA * area;

  weight ~ normal(mu, sigma);
  alpha ~ normal(4.5, 1);
  betaAF ~ normal(0, 5);
  betaG ~ normal(0, 5);
  betaA ~ normal(0, 5);
  sigma ~ uniform(0,5);
}
"
m3.2_data = list(
  N = nrow(data_assign3),
  avgfood = data_assign3$avgfood_z,
  groupsize = data_assign3$groupsize_z,
  area = data_assign3$area_z,
  weight = data_assign3$weight
)

m3.2_fit = stan(model_code = m3.2_draft, data = m3.2_data, iter = 2000, cores = 4, chains = 2)
plot(m3.2_fit, plotfun = "trace", pars = c("alpha", "betaG","betaAF","betaA","sigma"), inc_warmup = TRUE)
summary(m3.2_fit)
```
# 3-3 Is avgfood or area a better predictor of body weight?
```{r}
# m3.1 Fixed groupsize
m3.1_result = as.data.frame(m3.1_fit)
summary(data_assign3$avgfood)
new_avgfood = seq(0,1.5, length.out = 20)
m3.1.1_prediction = 
  tibble(avgfood = new_avgfood,
         predict_weight = purrr::map(new_avgfood - mean(data_assign3$avgfood),interval_est_2,
                                     alpha = m3.1_result$alpha,
                                     beta1 = m3.1_result$betaAF,
                                     beta2 = m3.1_result$betaG,
                                     x2 = 0,
                                     bound='')%>% unlist() %>% as.vector(),
         upper_95       = purrr::map(new_avgfood - mean(data_assign3$avgfood),interval_est_2,
                                     alpha = m3.1_result$alpha,
                                     beta1 = m3.1_result$betaAF,
                                     beta2 = m3.1_result$betaG,
                                     x2 = 0,
                                     bound ='upper',
                                     interval = 0.95)%>% unlist() %>% as.vector(),
         lower_95       = purrr::map(new_avgfood - mean(data_assign3$avgfood),interval_est_2,
                                     alpha = m3.1_result$alpha,
                                     beta1 = m3.1_result$betaAF,
                                    beta2 = m3.1_result$betaG,
                                     x2 = 0,
                                     bound='lower',
                                     interval = 0.95)%>% unlist() %>% as.vector()
         ) 
m3.1.1_plot = ggplot() + 
  geom_line(data = m3.1.1_prediction,
    aes(x = avgfood, y = predict_weight)) + 
  geom_ribbon(data = m3.1.1_prediction,
              aes(x = avgfood, ymin = lower_95, ymax = upper_95),
              alpha = .5)
```
Comparing the following 2 plots, the slope of the 2 models are both significant.
```{r}
m3.1.1_plot
m2.1.1_plot
```
Then, compare models with avgfood and area using RMSE.
```{r}
a = c()
for(i in c(1:116)){
  a = list.append(a,purrr::map(data_assign3$avgfood_z[i],interval_est_2,
                                     alpha = m3.1_result$alpha,
                                     beta1 = m3.1_result$betaAF,
                                     beta2 = m3.1_result$betaG,
                                     x2 = data_assign3$groupsize_z[i],
                                     bound='') %>% unlist() %>% as.vector())}
m3.1.1_RMSE = mean((a - data_assign3$weight) ** 2) ** 0.5
b = c()
for(i in c(1:116)){
  b = list.append(b,purrr::map(data_assign3$area_z[i],interval_est_2,
                                     alpha = m2.1_result$alpha,
                                     beta1 = m2.1_result$betaA,
                                     beta2 = m2.1_result$betaG,
                                     x2 = data_assign3$groupsize_z[i],
                                     bound='') %>% unlist() %>% as.vector())}
m2.1.1_RMSE = mean((b - data_assign3$weight) ** 2) ** 0.5
cat('RMSE of model with area =', m2.1.1_RMSE)
cat('RMSE of model with avgfood =',m3.1.1_RMSE)
```
From comparing the RMSEs, avgfood seems to be a better predictor in this case.

# 3-4 Why when both avgfood and area are in the same model, their effects are reduced?
From the following summary report, we can notice that avgfood and area both become insignificant.
The reason behind this is that the 2 variables is correlated with each other.
From the following scatter plot, we can notice that the 2 variables are positively correlated.
```{r}
summary(m3.2_fit)
plot(data_assign3$groupsize, data_assign3$area)
```
