---
title: "Assign6"
author: "Ryan Huang"
date: "5/21/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(tidyverse)
library(MASS)
library(gridExtra)
library(loo)
library(rethinking)
library(purrr)
map = purrr::map
select = dplyr::select
```
# Question 1
Model the relationship between density and percent cover, using a log- link
```{r}
data(salamanders, package = 'rethinking')
d = salamanders
# d$PCTCOVER = d$PCTCOVER / 100
d$PCTCOVER_log = ifelse(d$PCTCOVER==0, 0, log(d$PCTCOVER))
```
Stan Model
```{r}
m1.1='
data {
  int N;
  int SALAMAN[N]; 
  real PCTCOVER[N]; 
}
parameters {
  real a;
  real bp;
}
model {
  vector[N] lambda;
  a   ~ normal(0,100);
  bp  ~ normal(0,10);

  for(i in 1:N) lambda[i] = a + bp * (PCTCOVER[i]);
  SALAMAN ~ poisson_log(lambda);
}

generated quantities {
  vector[N] log_lik;
  vector[N] lambda;
  {
    for(i in 1:N) {
      lambda[i] = a + bp * (PCTCOVER[i]);
      log_lik[i] = poisson_log_lpmf(SALAMAN[i] | lambda[i]);
    }
  }
}
'
```
Data
```{r}
dat1.1 = list(
  N = nrow(d),
  SALAMAN = d$SALAMAN,
  PCTCOVER = d$PCTCOVER_log
)
```
Model fitting
```{r message=FALSE}
fit1.1 = stan(model_code = m1.1, data = dat1.1, cores = 4)
```
Prediction
```{r}
print(fit1.1, include = F, pars = 'log_lik', probs = c(.1, .5, .9))
post1.1 <- as.data.frame(fit1.1)

for( i in (1:nrow(d))){
  if(i == 1){
    pred = tibble(mean = exp(post1.1[,i+49]) %>% mean,
                  l_PI = exp(post1.1[,i+49]) %>% PI%>% .[1],
                  h_PI = exp(post1.1[,i+49]) %>% PI%>% .[2])
  } 
  else{
    pred =  tibble(mean = exp(post1.1[,i+49]) %>% mean,
                  l_PI = exp(post1.1[,i+49]) %>% PI%>% .[1],
                  h_PI = exp(post1.1[,i+49]) %>% PI%>% .[2]) %>% 
            rbind(pred, .)
  }
}
pred = pred %>% mutate(
  SALAMAN = d$SALAMAN,
  site = d$SITE,
  PCTCOVER = d$PCTCOVER_log
)

fig1.1 = pred %>% 
  ggplot() + 
  geom_point(aes(x=PCTCOVER, y=SALAMAN)) + 
  geom_ribbon(aes(x=PCTCOVER,
                  ymin = l_PI,
                  ymax = h_PI),
              alpha = 0.7)
fig1.1
```
From the above plot we can tell that the prediction result is poor. We can't infer the amount of salamanders by only including the percentage of PCTCOVER.

# Question 2
Improve the model by using other predictors, FORESTAGE

Model 2.1: lambda = a + bp * PCTCOVER + bf * FORESTAGE (w/o interaction)
```{r}
m2.1='
data {
  int N;
  int SALAMAN[N]; 
  real PCTCOVER[N]; 
  int FORESTAGE[N];
}
parameters {
  real a;
  real bp;
  real bf;
}
model {
  vector[N] lambda;
  a   ~ normal(0,100);
  bp  ~ normal(0,10);
  bf  ~ normal(0,10);

  for(i in 1:N) lambda[i] = a + bp * (PCTCOVER[i]) + bf * (FORESTAGE[i]);
  SALAMAN ~ poisson_log(lambda);
}

generated quantities {
  vector[N] log_lik;
  vector[N] lambda;
  {
    for(i in 1:N) {
      lambda[i] = a + bp * (PCTCOVER[i]) + bf * (FORESTAGE[i]);
      log_lik[i] = poisson_log_lpmf(SALAMAN[i] | lambda[i]);
    }
  }
}
'
```
Model 2.2: Add category variable cover = ifelse(PCTCOVER > 70, 1, 0)
w/ interaction between PCTCOVER and COVER
```{r}
m2.2='
data {
  int N;
  int SALAMAN[N]; 
  real PCTCOVER[N]; 
  int FORESTAGE[N];
  int COVER[N];
}
parameters {
  real a;
  real bp;
  real bf;
  real bc;
  real bpc;
}
model {
  vector[N] lambda;
  a   ~ normal(0,100);
  bp  ~ normal(0,10);
  bf  ~ normal(0,10);
  bc  ~ normal(0,10);
  bpc ~ normal(0,10);

  for(i in 1:N) lambda[i] = a + bp * (PCTCOVER[i]) + bf * (FORESTAGE[i]) + bc * (COVER[i]) + bpc * COVER[i] * PCTCOVER[i];
  SALAMAN ~ poisson_log(lambda);
}

generated quantities {
  vector[N] log_lik;
  vector[N] lambda;
  {
    for(i in 1:N) {
      lambda[i] = a + bp * (PCTCOVER[i]) + bf * (FORESTAGE[i]) + bc * (COVER[i]) + bpc * COVER[i] * PCTCOVER[i];
      log_lik[i] = poisson_log_lpmf(SALAMAN[i] | lambda[i]);
    }
  }
}
'
```
Model 2.3: no age, only cover, cover rate and cover * cover rate
```{r}
m2.3='
data {
  int N;
  int SALAMAN[N]; 
  real PCTCOVER[N]; 
  int COVER[N];
}
parameters {
  real a;
  real bp;
  real bc;
  real bpc;
}
model {
  vector[N] lambda;
  a   ~ normal(0,100);
  bp  ~ normal(0,10);
  bc  ~ normal(0,10);
  bpc ~ normal(0,10);

  for(i in 1:N) lambda[i] = a + bp * (PCTCOVER[i])  + bc * (COVER[i]) + bpc * COVER[i] * PCTCOVER[i];
  SALAMAN ~ poisson_log(lambda);
}

generated quantities {
  vector[N] log_lik;
  vector[N] lambda;
  {
    for(i in 1:N) {
      lambda[i] = a + bp * (PCTCOVER[i]) + bc * (COVER[i]) + bpc * COVER[i] * PCTCOVER[i];
      log_lik[i] = poisson_log_lpmf(SALAMAN[i] | lambda[i]);
    }
  }
}'
```
Model 2.5: cover, cover rate, cover * cover rate, 
```{r}
m2.5='
data {
  int N;
  int SALAMAN[N]; 
  real PCTCOVER[N]; 
  int FORESTAGE[N];
  int COVER[N];
}
parameters {
  real a;
  real bp;
  real bf;
  real bc;
  real bpc;
  real bfc;
  real bpf;
}
model {
  vector[N] lambda;
  a   ~ normal(0,100);
  bp  ~ normal(0,10);
  bf  ~ normal(0,10);
  bc  ~ normal(0,10);
  bpc ~ normal(0,10);
  bfc  ~ normal(0,10);
  bpf  ~ normal(0,10);

  for(i in 1:N) lambda[i] = a + bp * (PCTCOVER[i]) + bf * (FORESTAGE[i]) + bc * (COVER[i]) + bpc * COVER[i] * PCTCOVER[i] + bpf * PCTCOVER[i] * FORESTAGE[i] + bfc * FORESTAGE[i] * COVER[i];
  SALAMAN ~ poisson_log(lambda);
}

generated quantities {
  vector[N] log_lik;
  vector[N] lambda;
  {
    for(i in 1:N) {
      lambda[i] = a + bp * (PCTCOVER[i]) + bf * (FORESTAGE[i]) + bc * (COVER[i]) + bpc * COVER[i] * PCTCOVER[i] + bpf * PCTCOVER[i] * FORESTAGE[i] + bfc * FORESTAGE[i] * COVER[i];
      log_lik[i] = poisson_log_lpmf(SALAMAN[i] | lambda[i]);
    }
  }
}
'
```
Data2.1
```{r}
dat2.1 = list(
  N = nrow(d),
  SALAMAN   = d$SALAMAN,
  PCTCOVER  = d$PCTCOVER_log,
  FORESTAGE = d$FORESTAGE
)
```
Data2.2
```{r}
dat2.2 = list(
  N = nrow(d),
  SALAMAN   = d$SALAMAN,
  PCTCOVER  = d$PCTCOVER_log,
  FORESTAGE = d$FORESTAGE,
  COVER = ifelse(d$PCTCOVER>70,1,0)
)
```
Data2.3
```{r}
dat2.3 = list(
  N = nrow(d),
  SALAMAN   = d$SALAMAN,
  PCTCOVER  = d$PCTCOVER_log,
  COVER = ifelse(d$PCTCOVER>70,1,0)
)
```
Data2.4 Don't take log on PCTCOVER
```{r}
dat2.4 = list(
  N = nrow(d),
  SALAMAN   = d$SALAMAN,
  PCTCOVER  = d$PCTCOVER,
  COVER = ifelse(d$PCTCOVER>70,1,0)
)
```
Data2.5
```{r}
dat2.5 = list(
  N = nrow(d),
  SALAMAN   = d$SALAMAN,
  PCTCOVER  = d$PCTCOVER_log,
  FORESTAGE = d$FORESTAGE,
  COVER = ifelse(d$PCTCOVER>70,1,0)
)
```

Model fitting
```{r message=FALSE, warning=FALSE}
fit2.1 = stan(model_code = m2.1, data = dat2.1, cores = 4, iter = 4000)
fit2.2 = stan(model_code = m2.2, data = dat2.2, cores = 4, iter = 4000)
fit2.3 = stan(model_code = m2.3, data = dat2.3, cores = 4, iter = 4000)
fit2.4 = stan(model_code = m2.3, data = dat2.4, cores = 4, iter = 4000)
fit2.5 = stan(model_code = m2.5, data = dat2.5, cores = 4, iter = 4000)
```
Prediction of model 2.1
```{r}
print(fit2.1, include = F, pars = 'log_lik', probs = c(.1, .5, .9))
post2.1 <- as.data.frame(fit2.1)

for( i in (1:nrow(d))){
  if(i == 1){
    pred2 = tibble(mean = exp(post2.1[,i+50]) %>% mean,
                  l_PI = exp(post2.1[,i+50]) %>% PI%>% .[1],
                  h_PI = exp(post2.1[,i+50]) %>% PI%>% .[2])
  } 
  else{
    pred2 =  tibble(mean = exp(post2.1[,i+50]) %>% mean,
                  l_PI = exp(post2.1[,i+50]) %>% PI%>% .[1],
                  h_PI = exp(post2.1[,i+50]) %>% PI%>% .[2]) %>% 
            rbind(pred2, .)
  }
}
pred2 = pred2 %>% mutate(
  SALAMAN = d$SALAMAN,
  PCTCOVER = d$PCTCOVER_log,
  FORESTAGE = d$FORESTAGE
)
# Plot PCTCOVER
fig2.1.1 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=PCTCOVER, y=SALAMAN)) + 
  geom_ribbon(aes(x=PCTCOVER,
                  ymin = l_PI,
                  ymax = h_PI),
              alpha = 0.7)
fig2.1.1

# Plot FORESTAGE
fig2.1.2 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=FORESTAGE, y=SALAMAN)) + 
  geom_ribbon(aes(x=FORESTAGE,
                  ymin = l_PI,
                  ymax = h_PI),
              alpha = 0.7)
fig2.1.2
```
Prediction of model 2.2
```{r}
print(fit2.2, include = F, pars = 'log_lik', probs = c(.1, .5, .9))
post2.2 <- as.data.frame(fit2.2)
colnames(post2.2)

for( i in (1:nrow(d))){
  if(i == 1){
    pred2 = tibble(mean = exp(post2.2[,i+52]) %>% mean,
                  l_PI = exp(post2.2[,i+52]) %>% PI%>% .[1],
                  h_PI = exp(post2.2[,i+52]) %>% PI%>% .[2])
  } 
  else{
    pred2 =  tibble(mean = exp(post2.2[,i+52]) %>% mean,
                  l_PI = exp(post2.2[,i+52]) %>% PI%>% .[1],
                  h_PI = exp(post2.2[,i+52]) %>% PI%>% .[2]) %>% 
            rbind(pred2, .)
  }
}
pred2 = pred2 %>% mutate(
  SALAMAN = d$SALAMAN,
  PCTCOVER = d$PCTCOVER_log,
  FORESTAGE = d$FORESTAGE,
  COVER = as.factor(ifelse(d$PCTCOVER>70,1,0))
)
# Plot PCTCOVER
fig2.2.1 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=PCTCOVER, y=SALAMAN, color = COVER)) + 
  geom_ribbon(aes(x=PCTCOVER,
                  ymin = l_PI,
                  ymax = h_PI,
                  group = COVER,fill = COVER),
              alpha = 0.7)
fig2.2.1

# Plot FORESTAGE
fig2.2.2 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=FORESTAGE, y=SALAMAN, color = COVER)) + 
  geom_ribbon(aes(x=FORESTAGE,
                  ymin = l_PI,
                  ymax = h_PI,
                  group = COVER, fill=COVER),
              alpha = 0.7)
fig2.2.2
```
Prediction of model 2.3
```{r}
print(fit2.3, include = F, pars = 'log_lik', probs = c(.1, .5, .9))
post2.3 <- as.data.frame(fit2.3)

for( i in (1:nrow(d))){
  if(i == 1){
    pred2 = tibble(mean = exp(post2.3[,i+51]) %>% mean,
                  l_PI = exp(post2.3[,i+51]) %>% PI%>% .[1],
                  h_PI = exp(post2.3[,i+51]) %>% PI%>% .[2])
  } 
  else{
    pred2 =  tibble(mean = exp(post2.3[,i+51]) %>% mean,
                  l_PI = exp(post2.3[,i+51]) %>% PI%>% .[1],
                  h_PI = exp(post2.3[,i+51]) %>% PI%>% .[2]) %>% 
            rbind(pred2, .)
  }
}
pred2 = pred2 %>% mutate(
  SALAMAN = d$SALAMAN,
  PCTCOVER = d$PCTCOVER_log,
  FORESTAGE = d$FORESTAGE,
  COVER = as.factor(ifelse(d$PCTCOVER>70,1,0))
)
# Plot PCTCOVER
fig2.3.1 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=PCTCOVER, y=SALAMAN, color = COVER)) + 
  geom_ribbon(aes(x=PCTCOVER,
                  ymin = l_PI,
                  ymax = h_PI,
                  group = COVER,fill = COVER),
              alpha = 0.7)
fig2.3.1
```
Prediction of model 2.4
```{r}
print(fit2.4, include = F, pars = 'log_lik', probs = c(.1, .5, .9))
post2.4 <- as.data.frame(fit2.4)

for( i in (1:nrow(d))){
  if(i == 1){
    pred2 = tibble(mean = exp(post2.4[,i+51]) %>% mean,
                  l_PI = exp(post2.4[,i+51]) %>% PI%>% .[1],
                  h_PI = exp(post2.4[,i+51]) %>% PI%>% .[2])
  } 
  else{
    pred2 =  tibble(mean = exp(post2.4[,i+51]) %>% mean,
                  l_PI = exp(post2.4[,i+51]) %>% PI%>% .[1],
                  h_PI = exp(post2.4[,i+51]) %>% PI%>% .[2]) %>% 
            rbind(pred2, .)
  }
}
pred2 = pred2 %>% mutate(
  SALAMAN = d$SALAMAN,
  PCTCOVER = d$PCTCOVER,
  FORESTAGE = d$FORESTAGE,
  COVER = as.factor(ifelse(d$PCTCOVER>70,1,0))
)
# Plot PCTCOVER
fig2.4.1 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=PCTCOVER, y=SALAMAN, color = COVER)) + 
  geom_ribbon(aes(x=PCTCOVER,
                  ymin = l_PI,
                  ymax = h_PI,
                  group = COVER,fill = COVER),
              alpha = 0.7)
fig2.4.1
```
Prediction 2.5
```{r}
print(fit2.5, include = F, pars = 'log_lik', probs = c(.1, .5, .9))
post2.5 <- as.data.frame(fit2.5)

for( i in (1:nrow(d))){
  if(i == 1){
    pred2 = tibble(mean = exp(post2.5[,i+54]) %>% mean,
                  l_PI = exp(post2.5[,i+54]) %>% PI%>% .[1],
                  h_PI = exp(post2.5[,i+54]) %>% PI%>% .[2])
  } 
  else{
    pred2 =  tibble(mean = exp(post2.5[,i+54]) %>% mean,
                  l_PI = exp(post2.5[,i+54]) %>% PI%>% .[1],
                  h_PI = exp(post2.5[,i+54]) %>% PI%>% .[2]) %>% 
            rbind(pred2, .)
  }
}
pred2 = pred2 %>% mutate(
  SALAMAN = d$SALAMAN,
  PCTCOVER = d$PCTCOVER_log,
  FORESTAGE = d$FORESTAGE,
  COVER = as.factor(ifelse(d$PCTCOVER>70,1,0))
)
# Plot PCTCOVER
fig2.5.1 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=PCTCOVER, y=SALAMAN, color = COVER)) + 
  geom_ribbon(aes(x=PCTCOVER,
                  ymin = l_PI,
                  ymax = h_PI,
                  group = COVER,fill = COVER),
              alpha = 0.7)
fig2.5.1

# Plot FORESTAGE
fig2.5.2 = pred2 %>% 
  ggplot() + 
  geom_point(aes(x=FORESTAGE, y=SALAMAN, color = COVER)) + 
  geom_ribbon(aes(x=FORESTAGE,
                  ymin = l_PI,
                  ymax = h_PI,
                  group = COVER, fill=COVER),
              alpha = 0.7)
fig2.5.2
```
# Model Comparison
```{r}
fit_list <- list(fit1.1, fit2.1, fit2.2, fit2.3, fit2.4, fit2.5)
# extract log likelihoods
ll_list <- lapply(fit_list, extract_log_lik)
# exponentiate
exp_ll_list <- lapply(ll_list, exp)

waic_list <- list() 
for(i in 1:6) {
waic_list[[i]] <- waic(ll_list[[i]], r_eff = rel_n_eff_list[[i]], cores = 4)
}
names(waic_list) <- c('fit1.1', 'fit2.1', 'fit2.2', 'fit2.3', 'fit2.4', 'fit2.5')
loo::compare(x = waic_list)
```
Since model2.3 and model2.3 (include cover, cover % and interaction) are the 2 best models, we can tell that the covering percentage is more significant and we can tell that age of the forest isn't an important variable.  
The impacts of covering percentage will be different when covering rate is above or below 70%. 
