---
title: "Assign4"
author: "Ryan Huang"
date: "4/30/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
library(tidyverse)
library(rstan)
library(loo)
library(gridExtra)
options(mc.cores = 4)

```
## Question_1
```{r}
data('rugged', package = 'rethinking')
d <- rugged; rm(rugged)
d <-
  d %>%
  mutate(loggdp = log(rgdppc_2000))
d <-
  d %>% 
  filter(complete.cases(rgdppc_2000))
```

Question_1(a): Using rstan to fit just the interaction model
```{r results = "hide" }
m1.1 = "
data{
  int N;
  vector[N] loggdp;
  vector[N] rugged;
  vector[N] cont_africa;
}
parameters{
  real alpha;
  real bA;
  real bR;
  real bAR;
  real sigma;
}
model{
  // model
  vector[N] mu;
  for (i in 1:N){
      mu[i] = alpha + bA * cont_africa[i] + bR * rugged[i] + bAR * cont_africa[i] * rugged[i];
  }
  //prior
  alpha   ~ normal(8,100);
  bR  ~ normal(0,1);
  bA  ~ normal(0,1);
  bAR ~ normal(0,1);

  //likelihood
  loggdp ~ normal(mu,sigma);
}
generated quantities {
  vector[N] log_lik;
  {
    vector[N] mu;
    for(n in 1:N) {
    mu[n] = alpha + bA * cont_africa[n] + bR * rugged[n] + bAR * cont_africa[n] * rugged[n]; 
    log_lik[n] = normal_lpdf(loggdp[n] | mu[n], sigma);
    } 
  }
}
"
# Not for now
"
generated quantities {
  vector[N] log_lik;
  {
    vector[N] mu;
    for(n in 1:N) {
    mu[n] = alpha + bA * cont_africa[n] + bR * rugged[n] + bAR * cont_africa[n] * rugged[n]; 
    log_lik[n] = normal_lpdf(loggdp[n] | mu[n], sigma);
    } 
  }
}"

dat1.1 <- list(
  N = NROW(d),
  loggdp = d$loggdp,
  rugged = d$rugged, 
  cont_africa = d$cont_africa
)
fit1.1 = stan(model_code = m1.1, data = dat1.1, cores = 4, chains = 4, iter = 3000)
```
```{r}
print(fit1.1, probs = c(0.1, 0.5, 0.9), pars = c('alpha', 'bA', 'bR', 'bAR
', 'sigma'))
log_lik_1.1 <- extract_log_lik(fit1.1, merge_chains = FALSE)

```
Fit model w/o Seychelles
```{r results = "hide" }
d1.2 = d %>% filter(country != "Seychelles")
dat1.2 <- list(
  N = NROW(d1.2),
  loggdp = d1.2$loggdp,
  rugged = d1.2$rugged, 
  cont_africa = d1.2$cont_africa
)
fit1.2 = stan(model_code = m1.1, data = dat1.2, cores = 4, chains = 4, iter = 3000)
```
```{r}
print(fit1.2, probs = c(0.1, 0.5, 0.9), pars = c('alpha', 'bA', 'bR', 'bAR
', 'sigma'))
log_lik_1.2 <- extract_log_lik(fit1.2, merge_chains = FALSE)
```
```{r}
coeftab(fit1.1, fit1.2)
```

Comparing the 2 models, we can tell that the coefficients of the interaction term is decreased from 0.35 to 0.25 in the model w/o Seychelles.

Question_1(b): Plot the predictions of the interaction model, with and without Seychelles.
```{r}
# Plot w/ Seychelles.
post1.1 <- as.data.frame(fit1.1)
f_mu_1.1 <- function(rugged, cont_africa) with(post1.1,
alpha + bR * rugged + bAR * rugged * cont_africa + bA * cont_africa )

mu_1.1 <- mapply(f_mu_1.1, rugged = d$rugged, cont_africa = d$cont_africa)
mu_1.1_mean <- apply(mu_1.1, 2, mean)
mu_1.1_pi <- apply(mu_1.1, 2, rethinking::PI, prob = .97)
d_1.1 <- d %>%
  mutate(mu_mean = mu_1.1_mean, 
         mu_pi_l = mu_1.1_pi[1,], 
         mu_pi_h = mu_1.1_pi[2,],
         inAfrica = ifelse(cont_africa, 'African nations', 'Non-African
nations'))

fig1 = d_1.1 %>%
  ggplot(aes(x = rugged)) +
  geom_point(aes(rugged, loggdp, color = cont_africa), shape = 16) +
  theme(legend.position = '') +
  geom_line(aes(rugged, mu_mean)) +
  geom_ribbon(aes(x=rugged,ymin=mu_pi_l, ymax=mu_pi_h,fill = cont_africa),
              alpha = .1) +
  facet_wrap(~inAfrica) +  
  labs(x = 'Terrain Ruggedness Index', y = 'log GDP year 2000') +
  ggtitle("W/ Seychelles") +
  theme(plot.title = element_text(hjust = 0.5))

# Plot w/o Seychelles.
post1.2 <- as.data.frame(fit1.2)
f_mu_1.2 <- function(rugged, cont_africa) with(post1.2, alpha + bR * rugged + bAR * rugged * cont_africa + bA * cont_africa )

mu_1.2 <- mapply(f_mu_1.2, rugged = d$rugged, cont_africa = d$cont_africa)
mu_1.2_mean <- apply(mu_1.2, 2, mean)
mu_1.2_pi <- apply(mu_1.2, 2, rethinking::PI, prob = .97)
d_1.2 <- d %>%
  mutate(mu_mean = mu_1.2_mean, 
         mu_pi_l = mu_1.2_pi[1,], 
         mu_pi_h = mu_1.2_pi[2,],
         inAfrica = ifelse(cont_africa, 'African nations', 'Non-African
nations'))

fig2 = d_1.2 %>%
  ggplot(aes(x = rugged)) +
  geom_point(aes(rugged, loggdp, color = cont_africa), shape = 16) +
  theme(legend.position = '') +
  geom_line(aes(rugged, mu_mean)) +
  geom_ribbon(aes(x=rugged,ymin=mu_pi_l, ymax=mu_pi_h,fill = cont_africa),
              alpha = .1) +
  facet_wrap(~inAfrica) +  
  labs(x = 'Terrain Ruggedness Index', y = 'log GDP year 2000') +
  ggtitle("W/O Seychelles") +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(fig1, fig2, nrow=2)
```



Question_1(c): Conduct a model comparison analysis, using WAIC. Fit three models to the data without Seychelles. Then make model-averaged predictions of the 3 models.

Model_1: only consider rugged
```{r}
m3.1 = "
data{
  int N;
  vector[N] loggdp;
  vector[N] rugged;
}
parameters{
  real alpha;
  real bR;
real<lower=0, upper=10> sigma;
}
model{
  // model
  vector[N] mu;
  for (i in 1:N){
      mu[i] = alpha + bR * rugged[i];
  }
  // prior
  alpha ~ normal(8, 0.5);
  bR ~ normal(0, 0.25);
  
  // likelihood
  loggdp ~ normal(mu, sigma);
}
generated quantities{
  vector[N] log_lik;
  {
    vector[N] mu;
    for(i in 1:N){
      mu[i] = alpha + bR * rugged[i];
      log_lik[i] = normal_lpdf(loggdp[i] | mu[i], sigma); 
    }
  }
}
"
```
Model 2: Consider both rugged and cont_africa w/o interaction
```{r}
m3.2 = "
data{
  int N;
  vector[N] loggdp;
  vector[N] rugged;
  vector[N] cont_africa;
}
parameters{
  real alpha;
  real bR;
  real bA;
  real<lower=0, upper=5> sigma;
}
model{
  //model
  vector[N] mu;
  for (i in 1:N){
      mu[i] = alpha + bA * cont_africa[i] + bR * rugged[i];
  }
  //prior
  alpha ~ normal(8, 0.5);
  bR ~ normal(0, 0.5);
  bA ~ normal(0, 0.5);
  
  //likelihood
  loggdp ~ normal(mu, sigma);
}
generated quantities{
  vector[N] log_lik;
{
  vector[N] mu;
  for(i in 1:N){
    mu[i] = alpha + bR * rugged[i] + bA * cont_africa[i];
    log_lik[i] = normal_lpdf(loggdp[i] | mu[i], sigma);
    }
  }
}
"
```
Model 3: Consider R, A and the interaction.
```{r}
m3.3 = m1.1
```
Create data sets for 3 models
```{r results = "hide"}
dat3.1 = list(
  N = NROW(d1.2),
  loggdp = d1.2$loggdp,
  rugged = d1.2$rugged
)

dat3.2 = list(
  N = NROW(d1.2),
  loggdp = d1.2$loggdp,
  rugged = d1.2$rugged, 
  cont_africa = d1.2$cont_africa
)

dat3.3 = dat3.2
```
Fit model 3.1
```{r}
fit3.1 = stan(model_code = m3.1, data = dat3.1, cores = 4, chains = 4, iter = 3000)
```
Fit model 3.2
```{r output="hide"}
fit3.2 = stan(model_code = m3.2, data = dat3.2, cores = 4, chains = 4, iter = 3000)
```
Fit model 3.3
```{r results = "hide"}
fit3.3 = stan(model_code = m3.3, data = dat3.3, cores = 4, chains = 4, iter = 3000)
```
Model Comparison
```{r}
# extract log likelihood from the fitted model and use WAIC.
log_lik_3.1 = extract_log_lik(fit3.1, merge_chains = FALSE)
log_lik_3.2 = extract_log_lik(fit3.2, merge_chains = FALSE)
log_lik_3.3 = extract_log_lik(fit3.3, merge_chains = FALSE)
waic_3.1 = waic(log_lik_3.1)
waic_3.2 = waic(log_lik_3.2)
waic_3.3 = waic(log_lik_3.3)
mod_comp <- loo::compare(waic_3.1, waic_3.2, waic_3.3)
mod_comp
```
From the above comparison report, we can tell that model 3.3 with interaction term is still the best model regarding waic.

Plot the model-averaged predictions of the above 3 models.
```{r}
# get posterior of parameters
post3.1 <- as.data.frame(fit3.1) %>% select(alpha, bR)
post3.2 <- as.data.frame(fit3.2) %>% select(alpha, bR, bA)
post3.3 <- as.data.frame(fit3.3) %>% select(alpha, bR, bA, bAR)

# posterior predictors for model 3.1
f_mu_3.1 = function(rugged) with(post3.1, alpha + bR * rugged)
mu_3.1 = d1.2$rugged %>% purrr::map(f_mu_3.1)
mu_3.1_mean = mu_3.1 %>% purrr::map(mean) %>% purrr::flatten_dbl()
mu_3.1_pi = mu_3.1 %>% purrr::map(rethinking::PI, prob = .97) %>% unlist()

# posterior predictors for model 3.2
f_mu_3.2 = function(rugged, cont_africa) with(post3.2, alpha + bR * rugged + bA * cont_africa)
mu_3.2 = purrr::map2(d1.2$rugged,d1.2$cont_africa,f_mu_3.2)
mu_3.2_mean = mu_3.2 %>% purrr::map(mean) %>% purrr::flatten_dbl()
mu_3.2_pi = mu_3.2 %>% purrr::map(rethinking::PI, prob = .97) %>% unlist()

# posterior predictors for model 3.3
f_mu_3.3 = function(rugged,cont_africa) with(post3.3, alpha + bR * rugged + bAR * rugged * cont_africa + bA * cont_africa)
mu_3.3 = purrr::map2(d1.2$rugged, d1.2$cont_africa, f_mu_3.3)
mu_3.3_mean = mu_3.3 %>% purrr::map(mean) %>% purrr::flatten_dbl()
mu_3.3_pi = mu_3.3 %>% purrr::map(rethinking::PI, prob = .97) %>% unlist()

library(loo)
r_eff        <- relative_eff(exp(log_lik_3.1))
loo_3.1      <- loo(log_lik_3.1, r_eff = r_eff, cores = 2)
r_eff        <- relative_eff(exp(log_lik_3.2))
loo_3.2      <- loo(log_lik_3.2, r_eff = r_eff, cores = 2)
r_eff        <- relative_eff(exp(log_lik_3.3))
loo_3.3      <- loo(log_lik_3.3, r_eff = r_eff, cores = 2)

weights <- loo_model_weights(list(loo_3.1, loo_3.2, loo_3.3), method = "pseudobma", BB = F)

ensemble_mean <- weights[1] * mu_3.1_mean + weights[2] * mu_3.2_mean + weights[3] * mu_3.3_mean
ensemble_PI <- weights[1] * mu_3.1_pi + weights[2] * mu_3.2_pi + weights[3] * mu_3.3_pi

# Prepare data for plotting

d1.2 <- 
  d1.2 %>%
  mutate(ens_mean = ensemble_mean,
         ens_PI_l = ensemble_PI[seq(from=1, to=2*nrow(d1.2), by = 2)],
         ens_PI_h = ensemble_PI[seq(from=2, to=2*nrow(d1.2), by = 2)],
         inAfrica = ifelse(cont_africa, 'African nations', 'Non-African
nations'))

# Plotting the result
d1.2 %>%
  ggplot(aes(x = rugged)) +
  geom_point(aes(rugged, loggdp, color = cont_africa), shape = 16,) +
  theme(legend.position = '') +
  geom_line(aes(rugged, ens_mean)) +
  geom_ribbon(aes(x=rugged,ymin=ens_PI_l, ymax=ens_PI_h, fill = cont_africa),
              alpha = .1) +
  facet_grid(~inAfrica) +
  labs(x = 'Terrain Ruggedness Index', y = 'log GDP year 2000')

```

## Question_2
```{r}
data('nettle', package = 'rethinking')
d2 <- nettle; rm(nettle)
d2$lang.per.cap <- d2$num.lang / d2$k.pop
d2$log_lpc <- log(d2$lang.per.cap)
d2$log_area <- log(d2$area)
d2$log_area_c <- d2$log_area - mean(d2$log_area)
d2$mgs_c <- d2$mean.growing.season - mean(d2$mean.growing.season)
d2$sgs_c <- d2$sd.growing.season - mean(d2$sd.growing.season)

```
Question_2(a): Is log(lang.per_cap) positively associated with mean.growing.season? Consider log(area) as a covariate.
```{r output="hide"}
m2.1 = "
data{  
  int N;
  vector[N] log_lpc;
  vector[N] log_area;
  vector[N] mgs_c;
}
parameters{
  real alpha;
  real bA; //beta of area
  real bM; // beta of mean
  real<lower=0, upper=5> sigma;
}
model{
  // model
  vector[N] mu;
  for(i in 1:N){
    mu[i] = alpha + bA * log_area[i] + bM * mgs_c[i];
  }

  // Prior
  alpha ~ normal(-5, 1);
  bA ~ normal(0, 0.5);
  bM ~ normal(0, 0.25);

  // likelihood
  log_lpc ~ normal(mu, sigma);
}
"
dat2.1 = list(
  N = NROW(d2),
  log_lpc = d2$log_lpc,
  log_area = d2$log_area_c,
  mgs_c = d2$mgs_c
)
fit2.1 = stan(model_code = m2.1, data = dat2.1, cores = 4, chains = 4, iter = 3000)
```

```{r}
print(fit2.1)
```
Q2(b)
```{r output="hide"}
m2.2 = "
data{  
  int N;
  vector[N] log_lpc;
  vector[N] log_area;
  vector[N] sgs_c;
}
parameters{
  real alpha;
  real bA; //beta of area
  real bS; // beta of mean
  real<lower=0, upper=5> sigma;
}
model{
  // model
  vector[N] mu;
  for(i in 1:N){
    mu[i] = alpha + bA * log_area[i] + bS * sgs_c[i];
  }

  // Prior
  alpha ~ normal(-5, 1);
  bA ~ normal(0, 0.5);
  bS ~ normal(0, 0.25);

  // likelihood
  log_lpc ~ normal(mu, sigma);
}
"
dat2.2 = list(
  N = NROW(d2),
  log_lpc = d2$log_lpc,
  log_area = d2$log_area_c,
  sgs_c = d2$sgs_c
)
fit2.2 = stan(model_code = m2.2, data = dat2.2, cores = 4, chains = 4, iter = 3000)
```
```{r}
print(fit2.2)
```
Q2(c)
```{r output="hide"}
m2.3 = "
data{  
  int N;
  vector[N] log_lpc;
  vector[N] log_area;
  vector[N] sgs_c;
  vector[N] mgs_c;
}
parameters{
  real alpha;
  real bA; //beta of area
  real bS; // beta of std
  real bM; // beta of mean
  real bSM; // beta of interaction
  real<lower=0, upper=5> sigma;
}

model{
  // model
  vector[N] mu;
  for(i in 1:N){
    mu[i] = alpha + bA * log_area[i] + bS * sgs_c[i] + bM * mgs_c[i] + bSM * mgs_c[i] * sgs_c[i];
  }

  // Prior
  alpha ~ normal(-5, 1);
  bA ~ normal(0, 0.5);
  bS ~ normal(0, 0.25);
  bSM ~ normal(0, 0.5);

  // likelihood
  log_lpc ~ normal(mu, sigma);
}
"
dat2.3 = list(
  N = nrow(d2),
  log_lpc = d2$log_lpc,
  log_area = d2$log_area_c,
  sgs_c = d2$sgs_c,
  mgs_c = d2$mgs_c
)

fit2.3 = stan(model_code = m2.3, data = dat2.3, cores = 4, chains = 4, iter = 3000)
```
```{r}
print(fit2.3)
```

Plotting the prediction
```{r}
# Break the continous Xs
d2$mgsc.group <- cut(
  d2$mgs_c, 
  breaks = quantile(d2$mgs_c, probs = c(0, 1/3, 2/3, 1)),
  include.lowest = TRUE, 
  dig.lab = 2
)

d2$sgsc.group <- cut(
  d2$sgs_c, 
  breaks = quantile(d2$sgs_c, probs = c(0, 1/3, 2/3, 1)),
  include.lowest = TRUE, 
  dig.lab = 2
)

d2 = d2 %>% 
  group_by(mgsc.group) %>% 
  mutate(mgsc_gp_mean = mean(mgs_c)) %>% ungroup()
d2 = d2 %>% 
  group_by(sgsc.group) %>% 
  mutate(sgsc_gp_mean = mean(sgs_c)) %>% ungroup()

post2.3 = as.data.frame(fit2.3)
f_mu_2.3 = function(log_area,sgs_c,mgs_c) with(post2.3, alpha + bA * log_area + bS * sgs_c + bM * mgs_c + bSM * mgs_c * sgs_c)

# Grouped mgs_c
mu_2.3_A = purrr::pmap(list(d2$log_area_c, d2$sgs_c, d2$mgsc_gp_mean), f_mu_2.3)
mu_2.3_mean_A = mu_2.3_A %>% purrr::map(mean) %>% unlist()
mu_2.3_pi_A = mu_2.3_A %>% purrr::map(rethinking::PI, prob = .97) %>% unlist()

# Grouped sgs_c
mu_2.3_B = purrr::pmap(list(d2$log_area_c, d2$sgsc_gp_mean, d2$mgs_c), f_mu_2.3)
mu_2.3_mean_B = mu_2.3_B %>% purrr::map(mean) %>% unlist()
mu_2.3_pi_B = mu_2.3_B %>% purrr::map(rethinking::PI, prob = .97) %>% unlist()


figA = d2 %>% 
  mutate(
    mean = mu_2.3_mean_A,
    pi_l = mu_2.3_pi_A[seq(from=1, to=2*nrow(d2), by=2)],
    pi_h = mu_2.3_pi_A[seq(from=2, to=2*nrow(d2), by=2)]) %>%
  ggplot(aes(x = sgs_c)) +
  geom_point(aes(sgs_c, log_lpc), shape = 16,) +
  theme(legend.position = '') +
  geom_line(aes(sgs_c, mean)) +
  geom_ribbon(aes(x=sgs_c,ymin=pi_l, ymax=pi_h),
              alpha = .1)+
  facet_grid(~mgsc_gp_mean)

figB = d2 %>% 
  mutate(
    mean = mu_2.3_mean_B,
    pi_l = mu_2.3_pi_B[seq(from=1, to=2*nrow(d2), by=2)],
    pi_h = mu_2.3_pi_B[seq(from=2, to=2*nrow(d2), by=2)]) %>%
  ggplot(aes(x = mgs_c)) +
  geom_point(aes(mgs_c, log_lpc), shape = 16,) +
  theme(legend.position = '') +
  geom_line(aes(mgs_c, mean)) +
  geom_ribbon(aes(x=mgs_c,ymin=pi_l, ymax=pi_h),
              alpha = .1)+
  facet_grid(~sgsc_gp_mean)

grid.arrange(figA, figB, nrow=2)

```

