---
title: "Assign2"
author: "Ryan Huang"
date: "3/14/2019"
output: word_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rethinking)
library(rstan)
```
## Question 1
Re-center the weight data
```{r}
data("Howell1")
q1 <- Howell1 %>% mutate(weight.c = (weight - mean(weight)))

dat <- list(N = NROW(q1),
            height = q1$height,
            weight_c = q1$weight.c)

```

Fit the data to the model in class:
```{r}
q1_model = "
data {
  int<lower=1> N;
  vector[N] height;
  vector[N] weight_c;
}
parameters {
  real alpha;
  real beta;
  real sigma;
}
model {
  //Likelyhood
  vector[N] mu = alpha + beta * weight_c;
  
  //prior
  height ~ normal(mu, sigma);
  alpha ~ normal(178,20);
  beta ~ normal(0,10);
  sigma ~ uniform(0,50);
}
"
fit_q1 <- stan(model_code = q1_model, data = dat, cores = 4, chains = 2, iter = 1000)
```
The sampling results
```{r}
print(fit_q1, probs = c(0.10, 0.5, 0.9))
q1_post = as.data.frame(fit_q1)
p1 <- 
  ggplot() + 
  lims(y = c(130, 180)) + 
  geom_point(data = q1,
             aes(weight.c, height), 
             shape = 1, color = 'dodgerblue') +
  geom_abline(intercept = mean(q1_post$alpha), slope = mean(q1_post$beta))
```

The prediction result:
```{r}
interval_est = function(alpha,beta, x, bound){
  mu_at_x = alpha + beta * x
  if (bound == 'lower'){
    return(HPDI(mu_at_x)[1])
  } else if (bound == 'upper'){
    return(HPDI(mu_at_x)[2])
  } else {
    return(mean(mu_at_x))
  }
}

answer.sheet = 
  tibble(
    Individual = rep(1:5),
    Weight = c(46.95, 43.72, 64.78, 32.59, 54.63),
    )
answer.sheet = answer.sheet %>%
  mutate(
    expected_height = sapply(answer.sheet$Weight-mean(answer.sheet$Weight),
                      FUN=interval_est,
                      alpha=q1_post$alpha,
                      beta = q1_post$beta,
                      bound = ''),
    lower_89 = sapply(answer.sheet$Weight-mean(answer.sheet$Weight),
                      FUN=interval_est,
                      alpha=q1_post$alpha,
                      beta = q1_post$beta,
                      bound = 'lower'),
    upper_89 = sapply(answer.sheet$Weight-mean(answer.sheet$Weight),
                      FUN=interval_est,
                      alpha=q1_post$alpha,
                      beta = q1_post$beta, 
                      bound = 'upper')
  )
```
The final output
```{r}
answer.sheet
```


##Question 2

Filter the data of age under 18
```{r}
data('Howell1', package = 'rethinking')
q2_data = Howell1 %>% 
  filter( age < 18 )
q2_data = q2_data %>% mutate(weight.c = .$weight - mean(.$weight))
plot(q2_data$weight.c, q2_data$height)
```

Initialize Stan model
```{r}
q2_model = "
data {
  int N;
  vector[N] height;
  vector[N] weight_c;
}

parameters{
  real alpha;
  real beta;
  real sigma;
}
model{
//Likelyhood
  vector[N] mu = alpha + beta * weight_c;
  
  //prior
  height ~ normal(mu, sigma);
  alpha ~ normal(100,20);
  beta ~ normal(0,10);
  sigma ~ uniform(0,50);

}

"
```
Fit the data to the model
```{r}
dat = list(
        N = nrow(q2_data),
        height = q2_data$height,
        weight_c = q2_data$weight.c
)
fit_q2 <- stan(model_code = q2_model, data = dat, cores = 4, chains = 2, iter = 1000)

```
# MCMC result summary
The summary report tells us that the mean height of children under 18 is alpha = 108.34.
For every 10 units increase in weights, the model predicts the average height to increase by 27.2 units given other things are equal.
```{r}
print(fit_q2)
```



(b)Then we plot the raw data, the predicted regression line and 89% HPDI for the mean, and the 89% HPDI for predicted heights.
```{r}
q2_post = as.data.frame(fit_q2)    # the posterior from fit_q2 
f_mu <- function(x) q2_post$alpha + q2_post$beta * (x - mean(q2_data$weight)) 
# calculate mu for all combination of alpha, beta from post.

weight_new <- seq(0, 45)
mu <- # compute the 0.89 interval of regression line (predicted mu)
  sapply(weight_new, f_mu) %>%
  as_tibble() %>%
  rename_all(function(x) weight_new) %>%
  mutate(Iter = row_number()) %>%
  gather(weight, height, -Iter) %>%
  group_by(weight) %>%
  mutate(hpdi_l = HDInterval::hdi(height, credMass = 0.89)[1],
         hpdi_h = HDInterval::hdi(height, credMass = 0.89)[2]) %>%
  mutate(mu = mean(height)) %>%
  ungroup() %>%
  mutate(weight = as.integer(weight))


p2 <- ggplot() + xlim(0,50) +
  # plot the raw data
  geom_point(data = q2_data,
             aes(weight, height), shape = 1, color = 'dodgerblue') +
  # plot the CI
  geom_ribbon(data = mu,
              aes(x = weight, ymin = hpdi_l, ymax = hpdi_h),
              alpha = .5) +
  # plot the regression line
  # In the model x = weight.c, when plotting against weight, the intercept should be alpha - beta * mean(x)
  geom_abline(data = q2_post,
              aes(intercept = mean(alpha)- mean(q2_post$beta) * mean(q2_data$weight),
                  slope = mean(beta))) +
  labs(subtitle="HPDI Interval = 0.89")


#Add the Prediction interval
sim_ht <- 
  sapply(weight_new,
         function(x)
           rnorm(NROW(q2_post),
                 q2_post$alpha + q2_post$beta * (x - mean(q2_data$weight)),
                 q2_post$sigma)) %>%
  as_tibble() %>%
  rename_all(function(x) weight_new) %>%
  mutate(Iter = row_number()) %>%
  gather(weight, height, -Iter) %>%
  group_by(weight) %>%
  mutate(pi_l = rethinking::PI(height, prob = 0.89)[1],
         pi_h = rethinking::PI(height, prob = 0.89)[2]) %>%
  ungroup() %>%
  mutate(weight = as.integer(weight))

p2 + geom_ribbon(data = sim_ht,
                 mapping = aes(x=weight, ymin=pi_l, ymax=pi_h), alpha = .3) +
  labs(subtitle = 'Prediction Intervals = 0.89')
```
(c) From the plot in (b), we can notice that weight and height doesn't seem to have a linear relationship. When weights are small or large, the model tends to under estimate the height. However, when weights are around 15~30, the model usually over estimate the height. It seems that the prediction could improve if we take log on either the weights or the heights.

##Question 3

#(a)
Recode the stan model
```{r}
q3_model = "
data {
  int N;
  vector[N] height;
  vector[N] weight_log;
}
parameters{
  real alpha;
  real beta;
  real sigma;
}
model{
  //Likelyhood
  vector[N] mu = alpha + beta * weight_log;
  
  //prior
  height ~ normal(mu, sigma);
  alpha ~ normal(178,100);
  beta ~ normal(0,100);
  sigma ~ uniform(0,50);

}
"
```

Transform weight to log weight and 
```{r}
dat = list(
      N = nrow(Howell1),
      height = Howell1$height,
      weight_log = log(Howell1$weigh)
)
```
Fit the data to the model
```{r}
fit_q3 <- stan(model_code = q3_model, data = dat, cores = 4, chains = 2, iter = 1000)

print(fit_q3)
q3_post = as.data.frame(fit_q3)
(expected_mu_y = mean(q3_post$alpha) + mean(q3_post$beta) * log(mean(Howell1$weight)))

```
The alpha = -23.63 doesn't tell us much since we didn't centered the X.
After input the mean of the weight, the model predicts the mean height to be 144.39 units.
The summary report tells us that when weight increase by 1%, the height is expected to increase by beta/100 = 0.47 units.

#(b)
The raw data
```{r}
plot(height~weight, data=Howell1, col=col.alpha(rangi2, 0.4))
```
The regression result
```{r}
f_log_mu = function(x) mean(q3_post$alpha) + log(x) * mean(q3_post$beta)

mu = tibble(expected_mu = sapply(Howell1$weight, f_log_mu),
               weight = Howell1$weight)
p3 = ggplot() +
  # raw data
  geom_point(data = Howell1, aes(x=weight, y=height), alpha = 0.4, color = 'dodgerblue') +
  # regression line
  geom_line(data = mu, aes(x=weight, y=expected_mu)) + 
  labs(x = 'Weight', y = 'Height')
```
Add CI of mean
```{r}
# calculate mu for all combination of alpha, beta from post.
f_log_mu_matrix <- function(x) q3_post$alpha + q3_post$beta * log(x)

weight_new <- seq(5, 60)
mu_matrix <- # compute the 0.89 interval of regression line (predicted mu)
  sapply(weight_new, f_log_mu_matrix) %>%
  as_tibble() %>%
  rename_all(function(x) weight_new) %>%
  mutate(Iter = row_number()) %>%
  gather(weight, height, -Iter) %>%
  group_by(weight) %>%
  mutate(hpdi_l = HPDI(height)[1],
         hpdi_h = HPDI(height)[2]) %>%
  mutate(mu = mean(height)) %>%
  ungroup() %>%
  mutate(weight = as.integer(weight))

# Add to CI to p3
p3 = p3 +
  geom_ribbon(data = mu_matrix,
              aes(x = weight, ymin = hpdi_l, ymax = hpdi_h),
              alpha = .5)
```
Add PI
```{r}
#Add the Prediction interval
sim_ht <- 
  sapply(weight_new,
         function(x)
           rnorm(NROW(q3_post),
                 q3_post$alpha + q3_post$beta * log(x),
                 q3_post$sigma)) %>%
  as_tibble() %>%
  rename_all(function(x) weight_new) %>%
  mutate(Iter = row_number()) %>%
  gather(weight, height, -Iter) %>%
  group_by(weight) %>%
  mutate(pi_l = rethinking::PI(height, prob = 0.89)[1],
         pi_h = rethinking::PI(height, prob = 0.89)[2]) %>%
  ungroup() %>%
  mutate(weight = as.integer(weight))

p3 = p3 + geom_ribbon(data = sim_ht,
                 mapping = aes(x=weight, ymin=pi_l, ymax=pi_h), alpha = .3)
p3
```


